\textls[-0]{\thetitle}

### 2023-03-07 canonicitymodel.tex
Save.
Working on appendices.

\newcommand{\Glued}[1]{\ensuremath{{#1}^c}}
\newcommand{\GluedPi}[2]{\ensuremath{\Pi^c({#1},{#2})}}
\newcommand{\GSubstExt}[2]{\ensuremath{{#1},^c{#2}}}
\newcommand{\Gpair}[2]{\ensuremath{({#1},^c{#2})}}
\newcommand{\Gfst}[1]{\ensuremath{\texttt{fst}^c~{#1}}}
\newcommand{\Gsnd}[1]{\ensuremath{\texttt{snd}^c~{#1}}}
\newcommand{\Gsub}[2]{\ensuremath{{#1}\!\left[{#2}\right]^c}}
\newcommand{\Gapp}[1]{\ensuremath{\texttt{app}^c({#1})}}
\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^c)^{#1}}}
Once we have above model for MLTT, we can notice that we have a function $\Glued{\Pi}$ s.t. if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\GluedPi{A'}{B'} = \denotesC{\TyPi{A}{B}}$. Type-theoretical speaking, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model. Similarly, we can have $\Glued{\Sigma}$, $\Gpair{a}{b}$  and $\GSubstExt{\gm}{t}$  for dependent tuple type, depende tuple construction and substitution extension, and also all other constructions. These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.

@@

\begin{align*}
  \denotesC{\goodSig{\Gm}{\lsig}{n}} & \text{ is a list of 3-tuple of length $n$  } \\
  % & \text{ and we define  } {\goodSigC{\Gm}{\lsig}{n}} :\iff \denotesC{\goodSig{\Gm}{\lsig}{n}} \text{ is defined} \\ 
  \denotesC{\goodWSig{\Gm}{\wsig}{n}} &\text{ is a list of 2-tuple of length $n$ } \\
  % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
  & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
  \mathit{L}("nil") &= \denotesC{\top} \\
  \mathit{L}((A,s,T) :: tl)(\gm) &= {\mathit{L}(tl)}(\gm) \times \GluedPi{A}{T}(\gm) \quad \text{($\times$ is a special case for $\Glued{\Sigma}$)}  \\
  \mathit{P}("nil") &= \denotesC{\top}\\
  \mathit{P}((A,s,T) :: tl) &= \Glued{\Sigma}(\mathit{P}(tl), \Gsub{T}{\GSubstExt{\Glued{\pi_1}}{s}}) \quad \text{(doing substitution on $T$)} \\
  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \mathit{L}(\denotesC{\lsig}) \\
  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \mathit{P}(\denotesC{\lsig}) \\
  % 
  \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesC{A}, \denotesC{s}, \denotesC{T})::\denotesC{\lsig}\\ 
  \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[0] \quad \text{ take the first element in the tuple}\\ 
  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= ("hd"~\denotesC{\lsig})[1]\\ 
  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[2]\\
  \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm) &= (\denotesC{\ell}(\gm), \lambda u . t (\gm, u)) \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \\
  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\ell}}{\TyTkg{\lsig}}} &= \Gfst{\denotes{\ell}} \\
  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\ell}}{T}} &= \Gapp{\Gsnd{\denotes{\ell}}} \\
  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesC{A} , \denotesC{B}) :: \denotesC{\wsig}\\
  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[0]\\
  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[1]\\
  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}} &= \\
  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}}\\
  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \\
  \denotesC{\Recproj{j}{\lkg}} &= \\
  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}} &= \\
\end{align*}

>>>>

\newcommand{\Glued}[1]{\ensuremath{{#1}^c}}
\newcommand{\GluedPi}[2]{\ensuremath{\Pi^c({#1},{#2})}}
\newcommand{\GSubstExt}[2]{\ensuremath{{#1},^c{#2}}}
\newcommand{\Gpair}[2]{\ensuremath{({#1},^c{#2})}}
\newcommand{\Gfst}[1]{\ensuremath{\texttt{fst}^c~{#1}}}
\newcommand{\Gsnd}[1]{\ensuremath{\texttt{snd}^c~{#1}}}
\newcommand{\Gsub}[2]{\ensuremath{{#1}\!\left[{#2}\right]^c}}
\newcommand{\Gapp}[1]{\ensuremath{\texttt{app}^c({#1})}}
\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^c)^{#1}}}
\newcommand{\GLSigAdd}[3]{\ensuremath{\nu^{+c}({#1},{#2},{#3})}}
\newcommand{\GCaseSig}[3]{\ensuremath{\texttt{CaseTy}^c({#1},{#2},{#3})}}

Once we have above model for MLTT, we can notice that we have a function $\Glued{\Pi}$ s.t. if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\GluedPi{A'}{B'} = \denotesC{\TyPi{A}{B}}$. Type-theoretical speaking, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model. Similarly, we can have $\Glued{\Sigma}$, $\Gpair{a}{b}$  and $\GSubstExt{\gm}{t}$  for dependent tuple type, depende tuple construction and substitution extension, and also all other constructions. These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT. What's more, during the construction of the model for \TT, we will use $\Glued{(\cdot)}$ when we can.

@@

\begin{align*}
  \denotesC{\goodSig{\Gm}{\lsig}{n}} & \text{ is a list of 3-tuple of length $n$  } \\
  % & \text{ and we define  } {\goodSigC{\Gm}{\lsig}{n}} :\iff \denotesC{\goodSig{\Gm}{\lsig}{n}} \text{ is defined} \\ 
  \denotesC{\goodWSig{\Gm}{\wsig}{n}} &\text{ is a list of 2-tuple of length $n$ } \\
  % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
  & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
  \mathit{L}("nil") &= \denotesC{\top} \\
  \mathit{L}((A,s,T) :: tl)(\gm) &= {\mathit{L}(tl)}(\gm) \times \GluedPi{A}{T}(\gm) \quad \text{($\times$ is a special case for $\Glued{\Sigma}$)}  \\
  \mathit{P}("nil") &= \denotesC{\top}\\
  \mathit{P}((A,s,T) :: tl) &= \Glued{\Sigma}(\mathit{P}(tl), \Gsub{T}{\GSubstExt{\Glued{\pi_1}}{s}}) \quad \text{(doing substitution on $T$)} \\
  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \mathit{L}(\denotesC{\lsig}) \\
  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \mathit{P}(\denotesC{\lsig}) \\
  % 
  \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesC{A}, \denotesC{s}, \denotesC{T})::\denotesC{\lsig}\\ 
  \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[0] \quad \text{ take the first element in the tuple}\\ 
  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= ("hd"~\denotesC{\lsig})[1]\\ 
  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[2]\\
  \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm) &= (\denotesC{\lkg}(\gm), \lambda u . t (\gm, u)) \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \denotesC{\unit}
  \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gpair{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}{\Gsub{\Gsub{\denotesC{\lkgproj{2}{\lkg}}}{\GSubstExt{\Glued{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{(\SubstId)}}{\denotesC{\Tkg{\lkgproj{1}{\lkg}}}}}}
  \\ &\text{ when} \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\
  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyTkg{\lsig}}} &= \Gfst{\denotes{\lkg}} \\
  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\lkg}}{T}} &= \Gapp{\Gsnd{\denotes{\lkg}}} \\
  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesC{A} , \denotesC{B}) :: \denotesC{\wsig}\\
  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[0]\\
  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[1]\\
  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}} &= \\
  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
  % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\
  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= 
  {\GluedPi{\denotesC{A}}{\GluedPi {\GluedPi{\denotesC{B}}{\Gsub{\denotesC{R}}{\GSubstWeak{2}}}} {\Gsub{\denotesC{R}}{\GSubstWeak{2}}}}} \\
  % CsTy~\denotesC{A}~\denotesC{B}~\denotesC{R}\\
  \text{we define } \RecSig{}{} &\text{ by induction on the signature} \\
  RS~"nil"~R &= \denotesC{\LSigEmp} \\
  RS~((A, B) :: tl)~R &= \GLSigAdd{RS~tl~R}{\Glued{\pi_2}}{\GCaseSig{A}{B}{R}} \\
  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= RS~\denotesC{\tau}~\denotesC{R} \\
  \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg \\
  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}} &= \\
\end{align*}
